### 2014-06-10 old/vis14-usar/PacficVis Reviewer Rebuttal.txt
vmv14-usar: inital reduced version
git-svn-id: https://scivis.itn.liu.se/svn/publications/workinprogress/vmv14-usar@5439 cd0c9ba3-fe29-0410-b0c2-9a6eeba77867

We would like to thank the anonymous reviewers from PacificVis for their valuable
feedback that helped make the current revision much better and concise. In the following
we discuss how we addressed most of the reviewer`s comments.
The authors

Justification of design choices 
We included a requirements analysis and introduced the concept of sense-making and
decision-making theory to justify the decisions we made w.r.t. our system.

 Details of the implementation
  Performance and computation cost
We describe the details of the implementation in greater detail in the extended section 6.
The performance of the rendering, as well as the cost of the precomputation, is stated and
explained in this section as well.

 More information on the user feedback
We added a paragraph drawing conclusions from the feedback we received from the experts
during our qualitative evaluation.

As an application paper, a requirement analysis should be important. I
  hope the requirements and the supported tasks can be listed explicitly in
  the paper.
We adhered to this comment and included a requirement analysis and state the requirements
to our system in the introduction.

The workflow comparison in Section 3 is very important, since it tells
  the benefit of having the proposed system. However, it is not clear to me
  what are the differences of these two workflows. Maybe a list summarizing
  the differences would help.
We created a workflow figure (Figure 3) describing the differences between the current
workflow and our proposed solution.

For the path computation in Section 4.3, I wonder how many POIs can
  exist. If there are multiple POIs, how to plan the path with A*
  algorithm?
We added a discussion on this topic in Section 5.3

From Figure 1(bc), it seems that the point cloud rendering does not
  contain the ceiling/roof of the building. On the one hand, I suppose that
  the ceiling/roof data should be in the point cloud data, and should be
  important for spatial awareness. One the other hand, if the ceiling/roof
  is rendered, then the inner structure of the building will be blocked.
  This is not good. Therefore, I'm wondering how is ceiling/roof treated.
  Besides, a building usually has multiple floors. In this case, the
  occulsion problem will be serious in point cloud rendering. However, this
  is not mentioned in the paper.
This misunderstanding was clarified in Section 5.5 by adding a new part explaining the use
of interactive clip planes by the IC.

In the last paragraph of Section 4.3, it says "some might not arrive
  at the POI and are discarded". I wonder how is it possible. Given a start
  point and a POI, all paths calculated from A* algorithm should start from
  the start point and arrive at the POI. I suppose the only possibility of
  "not arrive" is that the path does not exist.
This point was clarified in Section 5.3 by explaining when a path does not exist.

In the first paragraph of Section 7, it says "to effectively annotate
  a three dimensional rendering of a filtered point cloud". I wonder how
  the point cloud is filtered.
The occupancy-based filtering is explained in greater detail in Sections 5.1 and 6.

The paper "prepared an expressive video showing the functionalities of
  out software".  The video helps much better understanding of the whole
  work. More information of this video should also be discussed in the
  paper, which is critical for readers to understand. 
We followed this suggestion and incorporated the video`s information into the paper in
various places.

First, many of the images chosen seem to be quite a poor representation
  of the system features. The teaser images were good; however, Figure 3
  left much to be desired.  Still images from the video could have been
  used to much greater effect (the video was very well done!).  In fact,
  the workflow in the video shows lots of really cool features which are
  poorly described in the paper.  
Figure 3 has been removed form the paper and the focus was shifted away from the
comparison with point-based rendering. We included more of the information contained in
the video into the paper.

Next, technical issues are glossed over.  Specifically Section 3 in the
  current workflow and visualization workflow, there is no sense of the
  amount of time it takes to generate these layouts, the co-registration,
  the fly-through, etc.  This was voiced by the users in the informal
  evaluation.  The runtime of the path computation should also be provided.
   If the goal of this system is to enable first responders, calculation
  time needs to be provided for assessment.
The timings for precomputations, renderings. path computations were added to the
Implementation section 6. The runtime for the co-registration is negligible as it is
performed during the real-time exploration (this is explained alongside the SLAM
reference)

So, from the abstract, the goal of this paper was to provide users
  information in a way that is easy to comprehend, and to "describe the
  design considerations for our system..." It seems likely that the
  consideration of path distance, inclination of walking surface, etc.
  would actually fit directly into a risk equation and provide some sort of
  metric (of course drilling down into multidimensional data is good and
  using PCP or SPLOMs can help).  However, from the informal feedback, the
  idea of assessing risk was not really discussed.  Also, in designing the
  scatterplot to handle four variables, using color and size as distinct
  variables can result in issues as color perception is directly related to
  the size of the area which is being colored.
This misunderstanding was hopefully removed by a more thorough description of the metric,
as this is the risk equation the reviewer is correctly talking about. The scatterplot
was replaced by a scatterplot matrix and the variable color and size were removed.

More details on the datatype for the point clouds should be given.  What
  is the size of the data, what fields are in the data (earlier there was a
  discussion of thermal for bodies and sound for heartbeats, is this
  captured?)
This point is now discussed in the Implementation section 6.

To summarize, this work is extremely interesting and the authors have
  made a large amount of good headway.  However, the paper needs a clear
  organization and goal.  If the goal is design constraints, it should
  focus more on this, describe how you can interact with the system,
  provide more details that are captured in the video.  If the paper is
  more focused on rendering of the building, then different tactics need to
  be chosen.  There is a lot of time spent on the point cloud rendering,
  but the technique used doesn't produce exceptionally great results. 
  However, that's not the crux of the work. Instead the authors wind up
  glossing over issues and leaving our crucial details.  For example, at
  the end of Section 4.5.2, the authors state "In this section we will
  describe the implementation details necessary to reproduce this work." 
  Then there are two sentences which basically states to use the OpenMP
  framework.  Is that all it takes for this?
The focus of the paper was removed from the point cloud rendering and shifted towards the
decision-making support of the system. Furthermore, the requested implementation section
was improved and describes the details.

Related work:
  The authors should also look at the following papers:
  William Ribarsky, Kalpathi R. Subramanian, Jianfei Liu, Onyewuchi
  Obirieze, Jack Guest: Mobile application for first response and emergency
  evacuation in Urban settings. COM.Geo 2010

SungYe Kim, Ross Maciejewski, Karl Ostmo, Edward J. Delp, Timothy F.
  Collins, David S.Ebert. Mobile Analytics for Emergency Response and
  Training. The Journal of Information Visualization, 7(1): 77-88, March
  2008.
Both papers were very relevant and were added to the discussion of related work.

The axes of the plots in Figure 6 are unreadable.  With overlaps in the
  profile plot, it is difficult to see which ones end earlier than others.
  Perhaps other visual encodings could have been used, or highlighting to
  help see the shortest paths?
The axes of the plots were redone. Furthermore, the plots are presented bigger to improve
readability more.

- Add a system overview, with the various components. Also, you should
  describe the implementation. In principle, the paper needs to be as
  reproducible as possible, and right now, it is not possible to know
  what's "under the hood".
A workflow figure was added to the paper, as well as an overview screenshot of the
complete system. The implementation is described in more detail in Section 6.