### 2014-06-10 old/eurovis14-usar/abstract.tex

\begin{abstract}
We propose a visualization system for incident commanders in urban search~\&~rescue scenarios that supports access path planning for post-disaster structures. Utilizing point cloud data acquired from unmanned robots, we combine visualization with visual analytics methods to allow for assessment of automatically generated access paths. Uncertainty in the data and a priori unknown information make fully automated systems impractical for this problem, thus requiring human interaction. To this end, we present a set of viable access paths, computed based on varying risk factors, to the incident commander in a three-dimensional environment combined with visual analysis tools to inspect paths in detail and make informed decisions. Based on this decision, a responder is guided along the path by the incident commander, who can interactively annotate and reevaluate the acquired point cloud. We describe design considerations for our system, its technical realization, and discuss the results of an expert evaluation.
\end{abstract}
\vspace*{-0.75cm}